Fast and Error-Adaptive Count-Distinct Sketch \\for Influence Maximization

### 2020-12-13 paper.tex
Update on Overleaf.

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{subfiles}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{subfig}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:~}}
\renewcommand{\algorithmicensure}{\textbf{Output:~}}
\newcommand\ovr[1]{\overrightarrow{#1}}

\usepackage{array}
\usepackage{color}
\usepackage{listings}
\usepackage{placeins}
\usepackage{enumitem}

\ifCLASSOPTIONcompsoc
  \usepackage[nocompress]{cite}
\else
  \usepackage{cite}
\fi

\ifCLASSINFOpdf
\else
\fi

\newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
pdftitle={Bare Demo of IEEEtran.cls for Computer Society Journals},%<!CHANGE!
pdfsubject={Typesetting},%<!CHANGE!
pdfauthor={Michael D. Shell},%<!CHANGE!
pdfkeywords={Computer Society, IEEEtran, journal, LaTeX, paper,
             template}}%<^!CHANGE!

@@

\title{Fast Influence Maximization Sketching\\ with Fusing and Vectorization }

\author{G\"{o}khan~G\"{o}kt\"{u}rk
        and~Kamer~Kaya% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem G. G\"{o}kt\"{u}rk and K. Kaya are with Computer Science and Engineering, Faculty of Engineering and Natural Sciences, Sabanci University, Istanbul, Turkey.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
}
%FIXME

\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Advanced Demo of IEEEtran.cls for IEEE Computer Society Journals}

\IEEEtitleabstractindextext{%
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}}

@@

Influence maximization is the problem of finding subset of $K$ vertices in a graph $G$ that has maximum reachability on some diffusion model $M$.

In addition to introducing the problem, Kempe et al.\cite{kempe2003maximizing} proved it to be NP-hard and provided a greedy Monte-Carlo approach that has a constant approximation for the optimal solution.

Many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, skim, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

\section{Notation and Background}\label{sec:background}

We denoted $G = (V,E)$ as a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations between the agents in $V$.
The neighborhood of a vertex $u \in V$ is denoted as $\Gamma_G(v) = \{v: (u,v) \in E\}$. 
A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. 
Diffusion probability on edge $(u, v) \in G$ is noted as $w_{u,v}$.
$w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(v)$   & Neighborhood of vertex $v$ in graph $G$\\
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & Influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
\hline\hline
        %$p$             & Sampling probability for edges\\
%        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $X_{r..n}$      & Slice of Vector X between r and r+n(not including)\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $B$             & Batch size\\
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & Sketch $j$th register for vertex $u$\\
        $\varsigma $    & Influence gained before sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after sketch build\\
        $\epsilon_g$    & Allowed global estimation error ratio\\
        $\epsilon_l$    & Allowed local estimation error\\ 
        $\epsilon_c$    & Allowed non-convergenced vertex ratio\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

@@

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.45\linewidth]{./images/ic-dif.png}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.45\linewidth]{./images/wc-dif.png}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph obtained from the directed one by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}
The influence spread function $\sigma_{G,M}(\cdot)$ computes the {\em expected} number of agents/nodes/vertices influenced~(activated) through a diffusion model $M$. For the sake of simplicity, we drop $M$ from the notation; in the rest of the text, $\sigma_{G}$ refers to $\sigma_{G,M}$. Some of the popular diffusion models for IM in the literature are {\em independent} and {\em weighted cascade}~(IC and WC),  and {\em linear threshold}~(LT)~\cite{kempe2003maximizing}.

\subsection{Sketch}\label{sec:sketch}
The count-distinct problem is the problem of finding the number of distinct elements in a stream with non-distinct elements. Computing reachability sets of a vertex is a similar problem; cardinality of all vertices visited is calculated while traversing sample sub-graphs from the given vertex. Exact cardinality calculations requires memory proportional to the cardinality.

Reachability set of a vertex is distinct union of all connected vertices' reachability sets. Many other Influence Maximization methods exploit this property in some degree.

The methods based on Reverse reachability and Bottom-Up Traversal, utilizes this property directly to merge reachability sets of connected vertices to estimate vertices influence.
MixGreedy method goes one step further; it utilizes that all vertices in a connected component that have the same reachability set in undirected graphs. So that, for a sample graph, all reachability sets can be found in single graph traversal.

@@

Count-distinct sketches have nice properties for such problems; Flajolet–Martin algorithm\cite{flajolet1985probabilistic} can estimate cardinality of distinct elements in reachability sets with constant number of registers($M$), and most importantly union of these sets can be done in constant number of operations.

Flajolet–Martin sketch only stores how rare elements are in a multi-set.Rarity of the elements are often computed by counting leading zeros in item's hash values.  The registers are initialized with zero values in begining. Items are added by storing longest leading zero count. Cardinality estimation can be done by taking 2 to the power of the stored value. Multiple registers and hash functions, $M[j]$ and $h_j$, are commonly used to reduce variance.
Adding item $x$ to register $M$, and its result $M'$ is shown in equation~\ref{eq:sketch-add}.

\begin{equation}
    \label{eq:sketch-add}
    M'[j] = max(M[j],clz(h_j(x)) ~ \forall j\in(0,J]
\end{equation} where $clz$ is count leading zeros function and $J$ is number of sketch registers.

@@

Average of longest running zeros can be used compute the cardinality and the result is divided to correction factor $\phi \approx 0.77351$ to correct hash collisions.
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $e$ distinct cardinality and $\bar{M}$ is M's registers mean.

@@

\subsection{SIMD}
{\em Single Instruction-Multiple Data} machines exploits data level parallelism to perform the same operation on multiple data points simultaneously. In this work, we utilized Advanced Vector Extensions (AVX2) instruction set. AVX2 works on 256-bit registers in multiple packed forms; 8x32 bit operations are used in fused sampling, where as sketch merge  performed with 32x8bit operations. This dichotomy between operations solved by processing in 32 simulations batches at a time.
Even-though, auto-vectorization is enabled, most of the operation mention are cannot be vectorized by the compiler. As far as we know, no compiler vectorizes chain conditional move or move-mask operations. Vectorization in the implementation is done using explicit intrinsics.

@@

For completeness, the intrinsics explicitly used in this paper are described in Table~\ref{tab:avx2-instructions}.

\begin{table}[!ht]
    \caption{AVX2 intrinsics used in the implementation.}
    \label{tab:avx2-instructions}
    \centering

    \begin{tabular}{|p{0.33\linewidth}|p{0.58\linewidth}|}
    \hline
    Intrinsic & Definition\\
    \hline
    {\tt \_mm256\_set1\_epi32} & Initializes 256-bit vector with scalar integer values. Doesn't map to any AVX instructions. \\
        % \hline
    {\tt \_mm256\_and\_si256} & Performs bitwise logical AND operation on 256-bit integer vectors. \\
            % \hline
    {\tt \_mm256\_xor\_si256} &  Performs bitwise logical XOR operation on 256-bit integer vectors.\\
            % \hline
    {\tt \_mm256\_cmpgt\_epi32} & Compares packed 8x 32-bit integers of two input vectors. \\
            % \hline
        % \hline
    {\tt \_mm256\_movemask\_ps} & Extracts the first bits of 8x 32-bit elements in a compact 8-bit format \\
            % \hline
    {\tt \_mm256\_blendv\_epi8} & Blends/selects byte elements of input vectors depending on the bits in a given mask vector.\\ 
             \hline         
    \end{tabular}
\end{table}{}

@@

The diffusion process works as follows; first, all vertices added to live vertices set $L$. Then, incoming edges $(u,v)$ for all live vertices are processed. For all simulation $u$ is live, the sketch of both sides are merged and set to $u$ registers. If any of the sketch-registers change, vertex $u$ added to the next iteration's live vertices set $L'$. After live edges are processed, if the next iteration's live vertices set is smaller than the convergence threshold $\epsilon_c$, the diffusion process concluded. If not, we swap $L$ and $L'$, clear $L'$, and then diffusion processed is repeated.

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{subfiles}
\usepackage{url}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{subfig}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:~}}
\renewcommand{\algorithmicensure}{\textbf{Output:~}}
\newcommand\ovr[1]{\overrightarrow{#1}}
\newcommand\myeq{\mkern1.5mu{=}\mkern1.5mu}
\usepackage{booktabs}
\usepackage{array}
\usepackage{color}
\usepackage{listings}
\usepackage{placeins}
\usepackage{enumitem}
%%%%%%
% Notes: 
% plot grouping/hue +
% table smallcaps +
% Time score memory center +
% Abstract
% Differences between infuser and us+
% time skecth diff between us +
% neden 0.005 +
% active vertex colors +
%%%%%
\ifCLASSOPTIONcompsoc
  \usepackage[nocompress]{cite}
\else
  \usepackage{cite}
\fi

\ifCLASSINFOpdf
\else
\fi
\newcommand\ggx[1]{\textcolor{blue}{#1}}

\newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
pdftitle={Bare Demo of IEEEtran.cls for Computer Society Journals},%<!CHANGE!
pdfsubject={Typesetting},%<!CHANGE!
pdfauthor={Gokhan Gokturk, Kamer Kaya},%<!CHANGE!
pdfkeywords={Influence Maximization, Graph Processing, Graph Sampling, Fused Sampling, Memory Access Regularization, Count-Distinct Sketch}}%<^!CHANGE!

\newcommand\acro{{\sc{HyperFuseR\xspace}\xspace}\xspace}
\usepackage{xcolor}
\newcommand\kktodo[1]{\textcolor{red}{#1}}

\newcommand\minspeedup{{{3.5\xspace}\xspace}\xspace}
\newcommand\maxspeedup{{{11\xspace}\xspace}\xspace}

\newcommand\maxspeedupTIM{{{1500\xspace}\xspace}\xspace}
\newcommand\maxspeedupIMM{{{27.87\xspace}\xspace}\xspace}
\newcommand\maxspeedupSKIM{{{11\xspace}\xspace}\xspace}

@@

\title{Fast and Error-Adaptive Count-Distinct Sketch
\\for Influence Maximization}

\author{G\"{o}khan~G\"{o}kt\"{u}rk
        and~Kamer~Kaya% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem G. G\"{o}kt\"{u}rk and K. Kaya are with Computer Science and Engineering, Faculty of Engineering and Natural Sciences, Sabanci University, TR~34956, Istanbul, Turkey.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
}
%FIXME

% \markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
% {Shell \MakeLowercase{\textit{et al.}}: Bare Advanced Demo of IEEEtran.cls for IEEE Computer Society Journals}

\IEEEtitleabstractindextext{%
\begin{abstract}
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation methods are frequently used for IM. In addition to high-quality yet expensive approximation algorithms, to meet with the scale of today's networks, lightweight, sketch-based approaches, which do not simulate the process exactly, have been proposed in the literature. As expected, the quality of the seed sets for the latter is inferior to those of the former. In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and the recently proposed hash-based, fused sampling. To estimate the vertex set sizes influenced throughout the simulations, we use per-vertex Flajolet-Martin sketches where each sketch register corresponds to a sample sub-graph. During the diffusion simulations, the reach set cardinalities of a single vertex are stored in adjacent registers in memory. This allows the proposed algorithm to find the number of influenced vertices in a single simulation step in different simulations at once. In addition, thanks to its efficiency and scalability of the our the parallel implementation can rebuild the sketches after observing large estimation errors. Our experimental results show that the proposed method finds high-quality seed sets  while being up-to $1.7\times$--$27.9\times$ faster on average than a state-of-the-art approximation algorithm. In addition, it is shown to be $3.5\times$--$11.1\times$ faster than a sketch-based approach while producing seed sets with $3\%$--$12\%$ better influence scores.
\end{abstract}
% Note that keywords are not normally used for peerreview papers.
}

@@

The study of efficient information/influence propagation in networks becomes an important area of research with several applications from many fields, such as viral marketing~\cite{leskovec2007dynamics, trusov2009effects}, social media analysis~\cite{zeng2010social, moreno2004dynamics}, and recommendation systems~\cite{lu2012recommender}.
As the study of these networks is imperative for educational, political, economic, and social purposes, a high-quality seed set to initiate the diffusion may have critical importance.
Furthermore, since the diffusion analysis may be time-critical, or increasing its coverage may be expensive in terms of computing resources and time, novel and efficient approaches to find good vertex sets that propagates the information/influence effectively is essential.

Influence maximization is the problem of finding a subset $S \subset V$ of $K$ vertices in a graph $G = (V, E)$ with the vertex set $V$ and edge set $E$ such that $S$ reaches the maximum reachability under some diffusion model $M$. In addition to introducing the problem, Kempe et al.~\cite{kempe2003maximizing} proved it to be NP-hard and provided a greedy Monte-Carlo approach that has a constant approximation for the optimal solution. This greedy approach is one of the most frequently applied algorithms for IM. The time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. Although they perform well in terms of seed-set quality, in consequence, the greedy Monte-Carlo solutions are impractical for real-life networks featuring millions of vertices. Due to this reason, many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, cohen2014sketch, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% However, these simulation-based, greedy algorithms provide the best possible approximation guarantees. Therefore they are considered as the gold standard for IM.

Simulating a greedy algorithm in parallel is an obvious and straightforward workaround to reduce the execution time of IM kernels and make them scalable for large-scale networks. However, for large networks, a greedy approach with a good approximation guarantee does not come cheap even if a large number of processing units/cores are available on networks with billions of vertices and edges. Following similar attempts in the literature, we propose a parallel, sketch-based greedy approach that approximates the Monte-Carlo processes. The proposed approach does not exactly count the number of influenced vertices. Instead, it leverages Count-Distinct sketches. 
in this work, we apply parallelism in multiple levels; thread-level and instruction-level parallelism. While modern compilers employ instruction-level parallelism effectively and automatically for applications with regular memory access patterns, graph kernels are usually harder to deal with. The irregularity of the neighbor IDs. ..... due to their irregular memory access patterns, it is not a straightforward task for graph processing kernels.  Besides, vectorization attempts on such kernels typically fail to provide substantial speed-up in performance.

In this work;
\begin{itemize}[leftmargin=*]
\item We propose, \acro\footnote{\url{https://github.com/ggokturk/infuser}}, an open-source, a blazing-fast, and accurate Influence Maximization algorithm. The proposed scheme samples the edges as they are traversed across several simulations. Thus, sampling, diffusion, and count-distinct processes are fused for all simulations.

\item Running concurrent simulations on per-vertex count-distinct sketches reduces the number of memory accesses. While traversing an edge, any number of diffusion simulations can be performed between vertices, using only a single integer (8-bit) per vertex for each simulation.

That is when the simulations traverse the same edge within close time intervals, the edge is accessed only once as long as the simulation timelines allow \acro to do so. The suggested approach, therefore, decreases the pressure on the memory management unit. In relation, for the cascade model, we employ a vector compute unit, near maximum efficiency to regularize memory accesses. 
%\item In addition to instruction-level parallelism, we parallelize \acro on multi-core architectures.

\item \acro{} can process very large graphs with millions of vertices and hundreds of millions of edges under a minute, without compromising the quality of results. For large graphs, due to its \ggx{race condition indifferent} nature, performance scales near linearly with the number of threads available. While processing the graph, only we few GBs of memory is used, mostly for storing the graph itself.

% \item \acro can be around $200000\times$ faster compared to the traditional greedy approaches. It is usable on large graphs that have been considered as {\em too large} in the literature. For instance, the new algorithm runs in $2.09$, $0.08$, $0.36$ seconds on networks {\tt Amazon}, {\tt NetHEP}, {\tt NetPhy} with 16 threads where the sequential baseline takes $141.3$, $259.1$ and $1725.2$ seconds, respectively. In fact, with a 302,400 seconds~(3.5 days) timeout, the sequential baseline can process only the above-mentioned 3 (out of 12) real-life graphs, having 1.2M, 58.9K and 231.5K edges. On the other graphs, the original algorithm cannot complete the simulations within the time limit. However, \acro{} completes all of 12 graphs around 1200 seconds in total, where the maximum is 654 seconds for the {\tt Orkut} network having 3.1M vertices and 117.2M edges.

\item We evaluate the performance, memory consumption, and influence score with a state-of-the-art influence maximization algorithms {\sc Skim}~\cite{cohen2014sketch}, {\sc Tim+}~\cite{tim}, {\sc Imm}~\cite{minutoli2019fast} in order to accurately position the performance of \acro{} in the IM literature. The experiments show that \acro is $\minspeedup\times$—$\maxspeedup\times$ faster than state-of-the-art while still being better in comparison in terms of scores, and using a comparable amount of memory.

\end{itemize}

The paper is organized as follows: 
In Section~\ref{sec:background}, we present 
the background on IM and introduce the mathematical notation. 
Section~\ref{sec:method} describes the proposed approach in detail.
In Section~\ref{sec:evaluation}, a thorough performance comparison over the traditional algorithms is provided by conducting experiments on various real-world datasets and influence settings. Besides, a comparison with the state-of-the-art from the literature is given.
Section~\ref{sec:relatedwork} presents a comparative overview of the existing work. Finally, Section~\ref{sec:conclusion} discusses future work and concludes the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Notation and Background}\label{sec:background}

Let $G = (V,E)$ be a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations between the agents in $V$.
The neighborhood of a vertex $u \in V$ is denoted as $\Gamma_{G}(v) = \{v: (u,v) \in E\}$. 
A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. 
Diffusion probability on edge $(u, v) \in G$ is noted as $w_{u,v}$.
$w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(v)$   & Neighborhood of incoming edges to vertex $v$ in graph $G$\\ %%FIXME
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & Influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
\hline\hline
        %$p$             & Sampling probability for edges\\
%        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $X_{r..n}$      & Slice of Vector X between r and r+n(not including)\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $B$             & Batch size\\
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & $j$th sketch register for vertex $u$\\
        $\varsigma $    & Influence gained before last sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after last sketch build\\
        $\epsilon_{global}$    & Global estimation error threshold\\
        $\epsilon_{local}$    & Local estimation error threshold\\ 
        $\epsilon_{conv}$    & Non-convergenced vertex threshold\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

@@

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.50\linewidth]{images/ic.pdf}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.50\linewidth]{images/wc.pdf}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph was obtained from the directed one by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}

The $\sigma_{G, M}(\cdot)$ influence spread function calculates the {\em estimated} number of agents / nodes / vertices affected~(activated) through a $M$ diffusion model. 
We eliminate $M$ from the notation for the sake of elegant simplicity; in the remainder of the text, $\sigma_{G}$ refers to $\sigma_{G, M}$.
In the literature, {\em independent} and {\em weighted cascade}~(IC and WC), and 
{\em linear threshold}~(LT)~\cite{kempe2003maximizing} are some of the widely recognized diffusion models for IM.

\begin{itemize}[leftmargin=*]
\item The {\bf Independent Cascade} process performs in iterations and activates the $v$ vertex if one of its (incoming) edges, $(u, v)$, is 
used during the iteration of diffusion, which has an expected probability of $w_{u, v}$, because $u$ was already influenced in the previous rounds. 
Activation probabilities are independent (from each other and previous activations) in the {\em independent} variant which we concentrate on in this paper, as in Figure~\ref{fig:ic}.
In theory, there can exist parallel and  independent $\{u, v\}$ edges in $E$. In practice, they are merged to a single $\{u,v\}$ edge via preprocessing.

\item The {\bf Weighted Cascade}  model is a variant of the independent cascade that uses the structural properties of vertices to set weights as shown in Figure~\ref{fig:wc}.
The method, as described in ~\cite{MixGreedy}, sets $w_{u, v} = 1 / d_v$ where $d_v$ is the number 
of incoming edges of $v$ for introducing weight values(which in the original graph is equal to $\Gamma G(v)$).
Therefore, if $v$ is triggered after the last round with $\ell$ neighbors, its activation probability in the new round is $1-( 1-1 / d_v)^\ell$.

%$ {|E^G_{u,v}|}/{|E^G_{u}|}$ where $|E^G_{u,v}|$ is the number of parallel $\{u, v\}$ edges and $|E^G_{u}|$ is the number of $\{u, .\}$ edges, i.e., all of $u$'s edges in the undirected graph.

\item{\bf Linear threshold} generalizes the cascade models and activates the $v$ vertex once the cumulative activation coming from the neighbors of $v$ exceeds the $\theta v$ threshold. 
All the $\{u, v\}$~(or $(u, v)$ in the directed graph) edges 
with active $u$ vertices are taken into account in the process. Vertex $v$ is activated~\cite{kempe2003maximizing} when the number of these edge weights reaches $\theta v$.  
Therefore, the cascade models are special variants of the linear threshold model with $\theta v = 0 $.

%Each edge $e=(s,v) \in E$ has a weight $w_{s,v}$. Each vertex $v$ has incoming edge list $Adj^{-1}(v)$ and $\sum_{s\in Adj^{-1}(v) }{w_{s,v}} \leq 1$. Also, each vertex $v$ has associated with a threshold $\theta_v$. Initially, $\theta$ is sampled uniformly, and seed set is set as activated. Then, until convergence is reached simulation ran to transfer $w_ij$ values from activated vertices to their neighborhoods, activating each vertex $v$ that reach $\theta_v$ value. Weights can be uniform i.e 0.1, 0.01, or 0.001 for all edges, as we observed in many work.

\item {\bf Triggering} proposed by Kempe et al.\cite{kempe2003maximizing} to generalize of previous models mentioned above. In the triggering model, each neighbor has a probability to influence vertex $v$. The diffusion process chooses a random subset of vertices called the "triggering set" to activate vertices at each instance.
\end{itemize}

The complexity analysis stays consistent for many diffusion models including {\em Independent Cascade}~(IC), {\em Linear Threshold}~(LT),  models. 
% Current studies focus on reducing the number of simulations run \cite{goyal2011celf++, goyal2011simpath}, sketch-based speedup methods\cite{SKIM}, and proxy methods that solve similar yet simpler problems \cite{kempe2003maximizing}. %FIXME too many methods, one french paper in python
%\cite{kempe2003maximizing} 
We concentrate on the independent cascade model in this paper, but the proposed methods are also relevant to the other models in the literature.
%  \[ \sigma(S^*)=\argmax_{S\subseteq V \land |S| \leq k } \sigma_{G}(S) \]

%is a simple diffusion model that Vertex v can be activated by any of incoming edges from activated vertices w with probability $p$. Given a seed set $S$ and influence probability $p$, cascade model starts from activated seed set and tries to diffuse until each of the out-going edges of active vertices are considered. The diffusion process considers all edges only once.

\subsection{Count-Distinct Sketch}\label{sec:sketch}
The count-distinct problem is the problem of finding the number of distinct elements in a stream with non-distinct elements. Computing reachability sets of a vertex is a similar problem; the cardinality of all vertices visited is calculated while traversing sample sub-graphs from the given vertex. Exact cardinality calculations require memory proportional to the cardinality.

The reachability set of a vertex is a distinct union of all connected vertices' reachability sets. Many other Influence Maximization methods exploit this property to some degree.

The methods based on Reverse reachability and Bottom-Up Traversal utilizes this property directly to merge reachability sets of connected vertices to estimate vertices influence.
MixGreedy method goes one step further; it utilizes that all vertices in a connected component that have the same reachability set in undirected graphs. So, for a sample graph, all reachability sets can be found in a single graph traversal.

@@

Count-distinct sketches have nice properties for such problems; Flajolet–Martin algorithm\cite{flajolet1985probabilistic} can estimate the cardinality of distinct elements in reachability sets with a constant number of registers($M$), and most importantly union of these sets can be done in a constant number of operations.

Flajolet–Martin sketch only stores how rare elements are in a multi-set.The rarity of the elements is often computed by counting leading zeros in the item's hash values.  The registers are initialized with zero values in beginning. Items are added by storing the longest leading zero count. Cardinality estimation can be done by taking 2 to the power of the stored value. Multiple registers and hash functions, $M[j]$ and $h_j$, are commonly used to reduce variance.
Adding item $x$ to register $M$, and its result $M'$ is shown in equation~\ref{eq:sketch-add}.

\begin{equation}
    \label{eq:sketch-add}
    M'[j] = max(M[j],clz(h_j(x)) ~ \forall j\in(0,J]
\end{equation} where $clz$ is count leading zeros function and $J$ is the number of sketch registers.

@@

Average of longest running zeros can be used compute the cardinality and the result is divided to correction factor $\phi \approx 0.77351$ to correct hash collisions.
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $e$ distinct cardinality and $\bar{M}$ is the mean of the registers.

@@

% \subsection{SIMD}
% {\em Single Instruction-Multiple Data} machines exploits data level parallelism to perform the same operation on multiple data points simultaneously. In this work, we utilized Advanced Vector Extensions (AVX2) instruction set. AVX2 works on 256-bit registers in multiple packed forms; 8x32 bit operations are used in fused sampling, where as sketch merge  performed with 32x8bit operations. This dichotomy between operations solved by processing in 32 simulations batches at a time.
% Even-though, auto-vectorization is enabled, most of the operation mention are cannot be vectorized by the compiler. As far as we know, no compiler vectorizes chain conditional move or move-mask operations. Vectorization in the implementation is done using explicit intrinsics.

@@

% For completeness, the intrinsics explicitly used in this paper are described in Table~\ref{tab:avx2-instructions}.

% \begin{table}[!ht]
%     \caption{AVX2 intrinsics used in the implementation.}
%     \label{tab:avx2-instructions}
%     \centering

%     \begin{tabular}{|p{0.33\linewidth}|p{0.58\linewidth}|}
%     \hline
%     Intrinsic & Definition\\
%     \hline
%     {\tt \_mm256\_set1\_epi32} & Initializes 256-bit vector with scalar integer values. Doesn't map to any AVX instructions. \\
%         % \hline
%     {\tt \_mm256\_and\_si256} & Performs bitwise logical AND operation on 256-bit integer vectors. \\
%             % \hline
%     {\tt \_mm256\_xor\_si256} &  Performs bitwise logical XOR operation on 256-bit integer vectors.\\
%             % \hline
%     {\tt \_mm256\_cmpgt\_epi32} & Compares packed 8x32-bit integers of two input vectors.\\
%     {\tt \_mm256\_max\_epi8} & Compares packed 32x8-bit integers of two input vectors and returns element-wise maximums. \\
%             % \hline
%         % \hline
%     {\tt \_mm256\_movemask\_ps} & Extracts the first bits of 8x 32-bit elements in a compact 8-bit format \\
%             % \hline
%     {\tt \_mm256\_blendv\_epi8} & Blends/selects byte elements of input vectors depending on the bits in a given mask vector.\\ 
%              \hline         
%     \end{tabular}
% \end{table}{}

@@

Second, the diffusion process is performed between vertex sketches. All vertices added to live vertices set $L$. Then, incoming edges $(u,v)$ for all live vertices are processed. For all simulation $u$ is live, the sketch of both sides are merged and set to $u$ registers. If any of the sketch-registers change, vertex $u$ is added to the next iteration's live vertices set $L'$. After live edges are processed, if the next iteration's live vertices set is smaller than the convergence threshold $\epsilon_c$, the diffusion process concluded. If not, we swap $L$ and $L'$, clear $L'$, and then diffusion processed is repeated.
